---
title: "Practice_Assignment_R_Programming"
author: "John Jurica"
date: "2023-09-26"
output: html_document
---



# https://github.com/rdpeng/practice_assignment/blob/master/practice_assignment.rmd 

#The first 2 lines of this command didn't work, so I am manually downloading and trying from there

dataset_url <- "http://s3.amazonaws.com/practice_assignment/diet_data.zip"
download.file(dataset_url, "diet_data.zip")

#The command below should still work. It is unzipping the fil and putting it in a folder called "diet_data inside the working directory file

```{r unzip data}
unzip("diet_data.zip", exdir = "diet_data")
```


#function shows the diet_data folder along with whatever else is in the working directory
```{r list files all files}
list.files()
```
# shows the five .csv files by only showing what's inside the working directory with the name "diet_data"

```{r list files within diet data}
list.files("diet_data")
```
# sets andy as a file we're reading... notice that we're still in a working directory a parent above where we're looking

```{r just read Andy.csv}
andy <- read.csv("diet_data/Andy.csv")
```

#gives us the header names

```{r get header names, echo = TRUE}
head(andy)
```

# see how many rows are in one of the columns
# Both the commands below work, one is looking at the dataset, which is andy. The first
# specifies that it's the column with the header Weight. The 2nd specifies that it's
# the column with a blank for the row cuz we're not looking at rows, then the 3rd column
#how does andy[ , 3] not grab the header too? Interesting question

```{r}
length(andy$Weight)
length(andy[ , 3])
```


Some other ways to get yourself some useful data

dim(andy)

#gets you some other helpful info

```{r}
str(andy)
summary(andy)
names(andy)
```

Let's take the output of `list.files()` and store it:
```{r}
files <- list.files("diet_data")
files
```



Let's take a quick look at John.csv:
```{r error=TRUE}
head(read.csv(files[3]))
```



Woah, what happened?  Well, John.csv is sitting inside the diet_data folder.  We just tried to run the equivalent of `read.csv("John.csv")` and R correctly told us that there isn't a file called John.csv in our working directory.  To fix this, we need to append the directory to the beginning of the file name.

One approach would be to use `paste()` or `sprintf()`.  However, if you go back to the help file for `list.files()`, you'll see that there is an argument called `full.names` that will append (technically prepend) the path to the file name for us.


files_full <- list.files("diet_data", full.names=TRUE)
files_full

Now we can read the individual files from here
head(read.csv(files_full[3]))


Success!  So what if we wanted to create one big data frame with everybody's data in it?  We'd do that with rbind and a loop.  Let's start with rbind:

THIS ASSUMES YOU HAVE ALREADY ADDED the file andy

```{r}
andy_david <- rbind(andy, read.csv(files_full[2]))
```

Success!  So what if we wanted to create one big data frame with everybody's data in it?  We'd do that with rbind and a loop.  Let's start with rbind:

```{r}
andy_david <- rbind(andy, read.csv(files_full[2]))
```

This works if you don't already have a dataframe for andy

david_john <- rbind(read.csv(files_full[2]), read.csv(files_full[3]))

Trying on my own to create a sum total here YIKES

total <- function()


How do I get these 2 to go together

Need an empty data frame
data <- data.frame()


for (i in 1:5) {
data <- rbind(data, read.csv(files_full[i])) 
}  

#This does not work... I think bevcause you cannot create the empty data frame in this setup

for (i in 1:5) {
  dat <- data.frame
        dat <- rbind(dat, read.csv(files_full[i]))
}

This turns it into a function and changes it to seq_along... In this situation, I am adding
print(head(dat) and print(nrow(dat)) becauses this allows me to check my work. I should get
the column headers with Name, Weight, and some other stuff. I should also get 150 rows from nrow

dat <- data.frame()

total <- function(y){
for (i in 1:5) {
        dat <- rbind(dat, read.csv(y[i]))
        
}
print(head(dat))
print(nrow(dat))
}

I got the expected values, so I can confirm that this has worked

Note that the function doesn't actually populate the dataframe named "dat"

attempting this to see if it worked
total(files_full)

Output confirms that it worked


This one also works, it just uses seq_along(y), which is better because
you don't have to input the exact number, but also could be worse if you're
using the numbers to specify which items in the list to grab

total <- function(y){
for (i in seq_along(y)) {
        dat <- rbind(dat, read.csv(y[i]))
}
dat
}

Cool.  We now have functions that are making a dataframe with all our stuff Out of curiousity, what would happen if we had put `dat <- data.frame()` inside of the loop?  Let's see:

```{r}
for (i in 1:5) {
        dat2 <- data.frame()
        dat2 <- rbind(dat2, read.csv(files_full[i]))
}
str(dat2)
head(dat2)
```

Because we put `dat2 <- data.frame()` inside of the loop, `dat2` is being rewritten with each pass of the loop.  So we only end up with the data from the last file in our list.



What if we wanted to make the number of csv files variable based on our input in a function

First attempt here didn't work, gave me 150 rows no matter what the argument was

total_test <- function(y,z){
for (z in seq_along(y)) {
        dat <- rbind(dat, read.csv(y[z]))
}
print(nrow(dat))
}


2nd attempt here, we are making the 2nd argument(z) in the function variable, whereas before 
it was hardcoded to grab everything in the first argument(y). This one worked!

total_test <- function(y,z){
for (i in z) {
        dat <- rbind(dat, read.csv(y[i]))
}
print(nrow(dat))
}


This was from an older way of doing it. It's nowhere near as good as the one above
I am going to skip this one and add the median part to the one above to finalize this


total_median_weight_day <- function(y,z){
     for (i in seq_along(y)) {
         dat <- rbind(dat, read.csv(y[i]))
     }
     hmm <- dat[which(dat[, "Day"] == z),]
     dis <- median(hmm$Weight, na.rm = TRUE)
     dis
 }


total_test <- function(y,z){
for (i in z) {
        dat <- rbind(dat, read.csv(y[i]))
}
print(nrow(dat))
}




total_data_day <- function(y,z){
     for (i in seq_along(y)) {
         dat <- rbind(dat, read.csv(y[i]))
     }
     hmm <- dat[which(dat[, "Day"] == z),]
     hmm
 }


Gets you the info for a specific day

dat_30 <- dat[which(dat[, "Day"] == 30),]












get_directory <- list.files("specdata", full.names=TRUE)


combine <- for (i in 1:332) {
dat <- rbind(dat, read.csv(get_directory[i]))
}


sub <- data.frame()

full_data_set <- for (i in seq_along(get_directory)) {
        sub <- rbind(sub, read.csv(get_directory[i]))
}



sub_23 <- sub[which(sub[, "ID"] == 23),]


pollutant <- function(directory) {
for (i in seq_along(directory)) {
hup <- rbind(hup, read.csv(directory[i]))
}


pollutant <- function(directory, pollutant, id) {
for (i in seq_along(directory)) {
hup <- rbind(hup, read.csv(directory[i]))
}
hmm <- hup[which(hup[, "ID"] %in% id),]
dis <- mean(hmm[,"pollutant"], na.RM = TRUE)
dis
}








Last try for the actual homework - I am trying to add an argument that lets me pick everything in a column.
And I am trying to change things to fit my parameters


total_mean_nitrate_id <- function(y,z){
     for (i in seq_along(y)) {
         dat <- rbind(dat, read.csv(y[i]))
     }
     hmm <- dat[which(dat[, "ID"] == z),]
     dis <- mean(hmm$nitrate, na.rm = TRUE)
     dis
 }


This is the answer for question 1

```{r} 

list_files_now <- function(directory,pollutant, id) {
     
a_directory <- list.files(directory, full.names=TRUE)

print(a_directory)

needed_files <- a_directory[id]

print(needed_files)

smoo <- data.frame()
for (i in seq_along(needed_files)) {
  smoo <- rbind(smoo, read.csv(needed_files[i]))
}
print(head(smoo))

final <- mean(smoo[,pollutant], na.rm = TRUE)

print(final)


}

list_files_now("specdata", "sulfate", 1:10)
```


This is also an answer for question 1
```{r}
 total_mean_pollutant_id <- function(directory,pollutant,id){
     #get a list of the directory
    get_directory <- list.files(directory, full.names=TRUE)
    
    #create an empty data frame
    dat <- data.frame()
    
    # get a subset of the needed stuff
    
    hmm <- get_directory[which(get_directory[, "ID"] %in% id),]
    
     # add needed files to dataframe
     for (i in seq_along(hmm)) {
         dat <- rbind(dat, read.csv(hmm[i]))
     }
     dis <- mean(dat[,pollutant], na.rm = TRUE)
     dis
 }
```

Trying again to figure out question 2

First part can be the same as question 1

get_id_and_nobs <- function(directory,id) {
     
a_directory <- list.files(directory, full.names=TRUE)

print(a_directory)

needed_files <- a_directory[id]

print(needed_files)

smoo <- data.frame()
for (i in seq_along(needed_files)) {
  smoo <- rbind(smoo, read.csv(needed_files[i]))
  final <- smoo %>%
        group_by(ID) %>%
        summarize(CompleteRows = sum(complete.cases(.)))

}
print(head(smoo))

final

}


get_id_and_nobs1 <- function(directory, id) {
    a_directory <- list.files(directory, full.names = TRUE)
    needed_files <- a_directory[id]
    
    # Initialize an empty data frame for results
    results <- data.frame()
    
    for (i in seq_along(needed_files)) {
        data <- read.csv(needed_files[i])
        complete_cases <- data %>% 
            group_by(ID) %>% 
            summarize(CompleteRows = sum(complete.cases(.)))
        results <- rbind(results, complete_cases)
    }
    
    return(results)
}

```{r}
get_id_and_nobs2 <- function(directory, id) {
    a_directory <- list.files(directory, full.names = TRUE)
    needed_files <- a_directory[id]
    
    # Initialize an empty data frame for results
    results <- tibble()
    
    for (i in seq_along(needed_files)) {
        data <- read.csv(needed_files[i])
        complete_cases <- data %>% 
            group_by(ID) %>% 
            summarize(nobs = sum(complete.cases(.)))
        results <- rbind(results, complete_cases)
    }
    
    return(results)
}
```


Part 3

get_cor_threshold <- function(directory, id, threshold) {
    a_directory <- list.files(directory, full.names = TRUE)
    needed_files <- a_directory[id]
    
    # Initialize an empty data frame for results
    results <- tibble()

    for (i in seq_along(needed_files)) {
        data <- read.csv(needed_files[i])
        complete_cases <- data %>%
            group_by(ID) %>%
            summarize(nobs = sum(complete.cases(.)))
        results <- rbind(results, complete_cases)
        
    }
    
    return(results)
}











total_obs_chat <- function(directory, id) {
    # Initialize an empty data frame to store the data
    dat <- NULL
    
    # Loop through each file in the directory and read the data
    for (i in seq_along(directory)) {
        data_file <- read.csv(directory[i])
        dat <- rbind(dat, data_file)
    }
    
    # Filter the data to include only rows with IDs in the specified 'id' vector
    hmm <- dat[dat$ID %in% id, ]
    hmm
    
    # Filter the data to include only complete cases for the "sulfate" column
    hma <- complete.cases(hmm)
    
    # Now 'hma' contains the data with the desired filters applied
    # You can perform further operations on 'hma' or return it
    
    # If you want to group the data by 'ID', you can use the 'aggregate' function:
    # aggregated_data <- aggregate(. ~ ID, data = hma, FUN = mean)
    
    # Return the filtered data or aggregated data
    return(hma)
}

